The truncated KdV framework for modeling anomalous waves induced by abrupt depth changes

### 2019-07-12 WavesData-2019/LatexNotes/notes.tex
NM

We introduce the following characteristic scales
\begin{align}
&\amp = \mean{\eta^2}^{1/2} 
&&\mbox{\em characteristic wave amplitude} \\
&\lam = \sqrt{g \depth} / \freqp
&&\mbox{\em characteristic wave length}
\end{align}
where $\mean{\cdot}$ indicates the mean of a quantity. Note that the characteristic wavelength $\lam = \lam(x)$ takes different values upstream and downstream of the ADC.
We introduce the following dimensionless variables
\begin{align}
&u = \eta/\amp
&&\mbox{\em dimensionless surface displacement} \\
&\tilde{x} = \frac{2 \pi \xi}{\lamfac \lam} = \frac{2 \pi (x-ct)}{\lamfac \lam}
&&\mbox{\em dimensionless position in moving frame} \\
& \tilde{t} = t \freqp
&&\mbox{\em dimensionless time} \\
\end{align}
Then the dimensionless KdV becomes
\begin{equation}
u_t + \frac{3 \pi}{\lamfac} \epsup \drat^{-1} u u_x + \frac{4 \pi^3}{3 \lamfac^3} \delup^2 \drat u_{xxx} = 0
\end{equation}
where $\epsup = \amp/\dup$, $\delup = \dup/\lamup = \sqrt{\dup \freqp^2/g}$, and $\drat = {\depth}/{\dup}$ is the local dimensionless depth.

\vsp{2}
Sidenote: If one simply uses the most naive scales, $u = \eta/\amp$, $\tilde{x} = \xi/\lam$, $\tilde{t} = t \freqp$, then the dimensionless KdV is the more standard one:
\begin{equation}
2 u_t + 3 \eps u u_x + \frac{\delta^2}{3} u_{xxx} = 0
\end{equation}
where $\eps = a/h$ and $\delta = h/\lam$.

\vsp{10}
Old notes here (I need to make sure everything is consistent and then remove this section below in favor of the newer, simpler writeup above).
The dimensionless KdV equation is
\begin{align}
\label{varKdV}
u_t + \ampp D^{-3/2} \, u u_x + \mu D^{1/2} \, u_{xxx} = 0
\end{align}
with the dimensionless parameters
\begin{align}
\label{ndKdV}
\ampp = \eps \delta^{-2} = \frac{ag}{h_0^2 f_p^2} , \qquad
\mu = \frac{4 \pi^2}{9 \lamfac^2}
\end{align}
Here, the displacement has been scaled on the characteristic wave amplitude, $u = \eta/a$, where $a = \etastd= \mean{\eta^2}^{1/2}$.
As a consequence, $u$ has fixed energy
\begin{equation}
E[u] = \frac{1}{2} \intt u^2 \dx = \frac{1}{2}
\end{equation}

In \eqref{ndKdV}, position $x$ has been scaled on the upstream characteristic wavelength $\lambda_0 = \sqrt{g h_0}/f_p$. Further, $\lamfac$ is the number of peak wavelengths in the periodic domain (perhaps we should set $\lamfac=1$, see note below).
Then the Hamiltonian is given by
\begin{align}
\label{Hamiltonian}
& \Hm = \ampp H_3 - \mu H_2 \\
& \Hp = \ampp D^{-3/2} H_3 - \mu D^{1/2} H_2
\end{align}
where
\begin{align}
&H_3 = \frac{1}{6} \intt u^3 \dx	\, , \\
&H_2 = \frac{1}{2} \intt u_x^2 \dx	\, .
\end{align}

>>>>

We introduce the following characteristic scales
\begin{align}
\label{ampdefn}
&\amp = \mean{\eta^2}^{1/2} 
&&\mbox{\em characteristic wave amplitude} \\
\label{lamdefn}
&\lam = \sqrt{g \depth} / \freqp
&&\mbox{\em characteristic wave length}
\end{align}
where $\mean{\cdot}$ indicates the mean of a quantity. Note that the characteristic wavelength $\lam = \lam(x)$ takes different values upstream and downstream of the ADC.
We introduce the following dimensionless variables
\begin{align}
&u = \eta/\amp
&&\mbox{\em dimensionless surface displacement} \\
&\tilde{x} = \frac{2 \pi \xi}{\lamfac \lam} = \frac{2 \pi (x-ct)}{\lamfac \lam}
&&\mbox{\em dimensionless position in moving frame} \\
& \tilde{t} = t \freqp
&&\mbox{\em dimensionless time} \\
\end{align}
We will consider the physical domain $\xi \in [0, \lamfac \lam]$, i.e.~an integer multiple of the characteristic wavelength. Then (after a shift) the dimensionless domain can be taken to be $\tilde{x} \in [-\pi,\pi]$.

The dimensionless KdV equation takes the form
\begin{equation}
u_t + \frac{3 \pi}{\lamfac} \epsup \drat^{-1} \, u u_x + \frac{4 \pi^3}{3 \lamfac^3} \delup^2 \drat \, u_{xxx} = 0
\qquad \text{for } x \in [-\pi,\pi]
\end{equation}
where dimensionless parameters are as follows
\begin{align}
&\epsup = \amp/\dup \, , \\
&\delup = \dup/\lamup = \sqrt{\dup \freqp^2/g} \, \\
&\drat(x) = {\depth(x)}/{\dup} \, .
\end{align}
$\drat = \drat(x)$ is the local dimensionless depth.

\begin{comment}
{\bf Sidenote}: In the case of constant depth, if one simply uses the most naive scales, $u = \eta/\amp$, $\tilde{x} = \xi/\lam$, $\tilde{t} = t \freqp$, then the dimensionless KdV is the more standard one:
\begin{equation}
2 u_t + 3 \eps u u_x + \frac{\delta^2}{3} u_{xxx} = 0
\end{equation}
where $\eps = a/h$ and $\delta = h/\lam$.
\end{comment}

\subsection{Hamiltonian analysis}

Note, that as a consequence of the definition of $\amp$ in \eqref{ampdefn}, we have that
\begin{equation}
E[u] = \frac{1}{2} \frac{1}{2 \pi} \int_{-\pi}^{\pi} u^2 \dx = \frac{1}{2}
\end{equation}

\vsp{10}

Then the Hamiltonian is given by
\begin{align}
\label{Hamiltonian}
& \Hm = \ampp H_3 - \mu H_2 \\
& \Hp = \ampp D^{-3/2} H_3 - \mu D^{1/2} H_2
\end{align}
where
\begin{align}
&H_3 = \frac{1}{6} \intt u^3 \dx	\, , \\
&H_2 = \frac{1}{2} \intt u_x^2 \dx	\, .
\end{align}
